---
title: "ST606 Thesis"
author: "Stephen Foley"
date: "3/30/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Packages
```{r}
library(tidyverse)
library(ggplot2)
library(TraMineR)
library(nnet)
library(factoextra)
library(cluster)
library(clValid)
library(WeightedCluster)
```

```{r}
data("biofam")



```

## Renaming columns
```{r}
biofam <- biofam %>% rename(
  nationality = nat_1_02
) %>% rename(
  language = plingu02
) %>% rename(
  religion = p02r01
) %>% rename(
  rel_participation = p02r04
) %>% rename(
  father_ss = cspfaj
) %>% rename(
  mother_ss = cspmoj
)

```

## Recoding states columns and retrieving the list of possible states
```{r}
recode1 <- function(x){
ifelse(x == 0, 'Parent',
       ifelse(x == 1, 'Left',
              ifelse(x == 2, 'Married',
                     ifelse(x == 3, 'Left+Marr',
                            ifelse(x == 4, 'Child',
                                   ifelse(x == 5, 'Left+Child',
                                          ifelse(x == 6, 'Left+Marr+Child',
                                                 ifelse(x == 7, 'Divorced', ''))))))))
}


biofam[,10:25] <- apply(biofam[,10:25], 2, recode1)








```


## Separating sequneces from covariates
```{r}
biofam$sex <- as.factor(biofam$sex)
biofam$nationality <- as.factor(biofam$nationality)
biofam$language <- as.factor(biofam$language)
biofam$religion <- as.factor(biofam$religion)
biofam$rel_participation <- as.factor(biofam$rel_participation)
biofam$father_ss <- as.factor(biofam$father_ss)
biofam$mother_ss <- as.factor(biofam$mother_ss)



biofam        <- list(covariates = biofam[c(2,6,8,9)],
                      sequences = biofam[,10:25], 
                      weights = biofam[,27])
biofam.cov      <- biofam$covariates

```


## Creating state sequence object
```{r}
biofam.alphabet <- c('Parent', 'Left', 'Married', 'Left+Marr', 'Child', 'Left+Child', 'Left+Marr+Child', 'Divorced')
biofam.labels <- c('Parent', 'Left', 'Married', 'Left+Marr', 'Child', 'Left+Child', 'Left+Marr+Child', 'Divorced')
biofam.scodes <- c('P', 'L', 'M', 'L+M', 'C', 'L+C', 'L+M+C', 'D')
biofam.seq <- seqdef(biofam$sequences, alphabet = biofam.alphabet, states = biofam.scodes,
                     labels = biofam.labels, xstep=8)

```


## Plotting the first 10 sequences, all the sequences and the most frequent sequences
```{r}
par(mfrow=c(2,2))

seqiplot(biofam.seq, with.legend = F, border = NA)
seqIplot(biofam.seq, sortv='from.start', with.legend=F)
seqfplot(biofam.seq, with.legend = F, border = NA)
seqlegend(biofam.seq)


```


## Plotting the state distributions, mean time spent in each state and the modal state sequences
```{r}
seqdplot(biofam.seq, with.legend=T, border=NA)

seqmtplot(biofam.seq, with.legend=F)

seqmsplot(biofam.seq, with.legend=F, border=NA)

```

## Estimating substitution cost
```{r}
subt <- seqsubm(biofam.seq, method='TRATE')
subc <- seqsubm(biofam.seq, method='CONSTANT')
subf <- seqsubm(biofam.seq, method='FUTURE')
subi <- seqsubm(biofam.seq, method='INDELS')
subig <- seqsubm(biofam.seq, method='INDELSLOG')



```


## Constructing pairwise matrices using different distances measures
```{r warning=F, message=F}
biofam.OM <- seqdist(biofam.seq, method = 'OM', indel=1, sm=subt)
biofam.DHD <- seqdist(biofam.seq, method='DHD')
biofam.OMl <- seqdist(biofam.seq, method='OMloc', indel=1, sm=subt)
biofam.OMsl <- seqdist(biofam.seq, method='OMslen', indel=1, sm=subt)
biofam.OMsp <- seqdist(biofam.seq, method='OMspell', indel=1, sm=subt)
biofam.H <- seqdist(biofam.seq, method='HAM')
biofam.CH <- seqdist(biofam.seq, method='CHI2')
biofam.EU <- seqdist(biofam.seq, method='EUCLID')
biofam.LS <- seqdist(biofam.seq, method='LCS')
biofam.LP <- seqdist(biofam.seq, method='LCP')
biofam.RL <- seqdist(biofam.seq, method='RLCP')
biofam.N <- seqdist(biofam.seq, method='NMS')
biofam.NT <- seqdist(biofam.seq, method='NMSMST')
biofam.SV <- seqdist(biofam.seq, method='SVRspell')


```


## Hierachical clustering methods
### Using OM matrix
```{r}
hier.OM.W <- cluster::agnes(biofam.OM, diss=T, method='ward')
hier.OM.A <- cluster::agnes(biofam.OM, diss=T, method='average')
hier.OM.S <- cluster::agnes(biofam.OM, diss=T, method='single')
hier.OM.C <- cluster::agnes(biofam.OM, diss=T, method='complete')
hier.OM.WE <- cluster::agnes(biofam.OM, diss=T, method='weighted')
hier.OM.GA <- cluster::agnes(biofam.OM, diss=T, method='gaverage')


hier.OM.W1 <- list()
hier.OM.A1 <- list()
hier.OM.S1 <- list()
hier.OM.C1 <- list()
hier.OM.WE1 <- list()
hier.OM.GA1 <- list()
for (k in 1:20) {
hier.OM.W1[[k]] <- cutree(hier.OM.W, k=k)
hier.OM.A1[[k]] <- cutree(hier.OM.A, k=k)
hier.OM.S1[[k]] <- cutree(hier.OM.S, k=k)
hier.OM.C1[[k]] <- cutree(hier.OM.C, k=k)
hier.OM.WE1[[k]] <- cutree(hier.OM.WE, k=k)
hier.OM.GA1[[k]] <- cutree(hier.OM.GA, k=k)
}

```


### Using DHD matrix
```{r}
hier.DHD.W <- cluster::agnes(biofam.DHD, diss=T, method='ward')
hier.DHD.A <- cluster::agnes(biofam.DHD, diss=T, method='average')
hier.DHD.S <- cluster::agnes(biofam.DHD, diss=T, method='single')
hier.DHD.C <- cluster::agnes(biofam.DHD, diss=T, method='complete')
hier.DHD.WE <- cluster::agnes(biofam.DHD, diss=T, method='weighted')
hier.DHD.GA <- cluster::agnes(biofam.DHD, diss=T, method='gaverage')




hier.DHD.W1 <- list()
hier.DHD.A1 <- list()
hier.DHD.S1 <- list()
hier.DHD.C1 <- list()
hier.DHD.WE1 <- list()
hier.DHD.GA1 <- list()
for (k in 1:20) {
hier.DHD.W1[[k]] <- cutree(hier.DHD.W, k=k)
hier.DHD.A1[[k]] <- cutree(hier.DHD.A, k=k)
hier.DHD.S1[[k]] <- cutree(hier.DHD.S, k=k)
hier.DHD.C1[[k]] <- cutree(hier.DHD.C, k=k)
hier.DHD.WE1[[k]] <- cutree(hier.DHD.WE, k=k)
hier.DHD.GA1[[k]] <- cutree(hier.DHD.GA, k=k)
}

```

### Using OMloc matrix
```{r}
hier.OMl.W <- cluster::agnes(biofam.OMl, diss=T, method='ward')
hier.OMl.A <- cluster::agnes(biofam.OMl, diss=T, method='average')
hier.OMl.S <- cluster::agnes(biofam.OMl, diss=T, method='single')
hier.OMl.C <- cluster::agnes(biofam.OMl, diss=T, method='complete')
hier.OMl.WE <- cluster::agnes(biofam.OMl, diss=T, method='weighted')
hier.OMl.GA <- cluster::agnes(biofam.OMl, diss=T, method='gaverage')


hier.OMl.W1 <- list()
hier.OMl.A1 <- list()
hier.OMl.S1 <- list()
hier.OMl.C1 <- list()
hier.OMl.WE1 <- list()
hier.OMl.GA1 <- list()
for (k in 1:20) {
hier.OMl.W1[[k]] <- cutree(hier.OMl.W, k=k)
hier.OMl.A1[[k]] <- cutree(hier.OMl.A, k=k)
hier.OMl.S1[[k]] <- cutree(hier.OMl.S, k=k)
hier.OMl.C1[[k]] <- cutree(hier.OMl.C, k=k)
hier.OMl.WE1[[k]] <- cutree(hier.OMl.WE, k=k)
hier.OMl.GA1[[k]] <- cutree(hier.OMl.GA, k=k)
}

```

### Using OMslen matrix
```{r}
hier.OMsl.W <- cluster::agnes(biofam.OMsl, diss=T, method='ward')
hier.OMsl.A <- cluster::agnes(biofam.OMsl, diss=T, method='average')
hier.OMsl.S <- cluster::agnes(biofam.OMsl, diss=T, method='single')
hier.OMsl.C <- cluster::agnes(biofam.OMsl, diss=T, method='complete')
hier.OMsl.WE <- cluster::agnes(biofam.OMsl, diss=T, method='weighted')
hier.OMsl.GA <- cluster::agnes(biofam.OMsl, diss=T, method='gaverage')


hier.OMsl.W1 <- list()
hier.OMsl.A1 <- list()
hier.OMsl.S1 <- list()
hier.OMsl.C1 <- list()
hier.OMsl.WE1 <- list()
hier.OMsl.GA1 <- list()
for (k in 1:20) {
hier.OMsl.W1[[k]] <- cutree(hier.OMsl.W, k=k)
hier.OMsl.A1[[k]] <- cutree(hier.OMsl.A, k=k)
hier.OMsl.S1[[k]] <- cutree(hier.OMsl.S, k=k)
hier.OMsl.C1[[k]] <- cutree(hier.OMsl.C, k=k)
hier.OMsl.WE1[[k]] <- cutree(hier.OMsl.WE, k=k)
hier.OMsl.GA1[[k]] <- cutree(hier.OMsl.GA, k=k)
}

```

### Using OMspell matrix
```{r}
hier.OMsp.W <- cluster::agnes(biofam.OMsp, diss=T, method='ward')
hier.OMsp.A <- cluster::agnes(biofam.OMsp, diss=T, method='average')
hier.OMsp.S <- cluster::agnes(biofam.OMsp, diss=T, method='single')
hier.OMsp.C <- cluster::agnes(biofam.OMsp, diss=T, method='complete')
hier.OMsp.WE <- cluster::agnes(biofam.OMsp, diss=T, method='weighted')
hier.OMsp.GA <- cluster::agnes(biofam.OMsp, diss=T, method='gaverage')


hier.OMsp.W1 <- list()
hier.OMsp.A1 <- list()
hier.OMsp.S1 <- list()
hier.OMsp.C1 <- list()
hier.OMsp.WE1 <- list()
hier.OMsp.GA1 <- list()
for (k in 1:20) {
hier.OMsp.W1[[k]] <- cutree(hier.OMsp.W, k=k)
hier.OMsp.A1[[k]] <- cutree(hier.OMsp.A, k=k)
hier.OMsp.S1[[k]] <- cutree(hier.OMsp.S, k=k)
hier.OMsp.C1[[k]] <- cutree(hier.OMsp.C, k=k)
hier.OMsp.WE1[[k]] <- cutree(hier.OMsp.WE, k=k)
hier.OMsp.GA1[[k]] <- cutree(hier.OMsp.GA, k=k)
}

```


### Using OMspell matrix
```{r}
hier.OMsp.W <- cluster::agnes(biofam.OMsp, diss=T, method='ward')
hier.OMsp.A <- cluster::agnes(biofam.OMsp, diss=T, method='average')
hier.OMsp.S <- cluster::agnes(biofam.OMsp, diss=T, method='single')
hier.OMsp.C <- cluster::agnes(biofam.OMsp, diss=T, method='complete')
hier.OMsp.WE <- cluster::agnes(biofam.OMsp, diss=T, method='weighted')
hier.OMsp.GA <- cluster::agnes(biofam.OMsp, diss=T, method='gaverage')


hier.OMsp.W1 <- list()
hier.OMsp.A1 <- list()
hier.OMsp.S1 <- list()
hier.OMsp.C1 <- list()
hier.OMsp.WE1 <- list()
hier.OMsp.GA1 <- list()
for (k in 1:20) {
hier.OMsp.W1[[k]] <- cutree(hier.OMsp.W, k=k)
hier.OMsp.A1[[k]] <- cutree(hier.OMsp.A, k=k)
hier.OMsp.S1[[k]] <- cutree(hier.OMsp.S, k=k)
hier.OMsp.C1[[k]] <- cutree(hier.OMsp.C, k=k)
hier.OMsp.WE1[[k]] <- cutree(hier.OMsp.WE, k=k)
hier.OMsp.GA1[[k]] <- cutree(hier.OMsp.GA, k=k)
}

```


### Using HAM matrix
```{r}
hier.OMsp.W <- cluster::agnes(biofam.OMsp, diss=T, method='ward')
hier.OMsp.A <- cluster::agnes(biofam.OMsp, diss=T, method='average')
hier.OMsp.S <- cluster::agnes(biofam.OMsp, diss=T, method='single')
hier.OMsp.C <- cluster::agnes(biofam.OMsp, diss=T, method='complete')
hier.OMsp.WE <- cluster::agnes(biofam.OMsp, diss=T, method='weighted')
hier.OMsp.GA <- cluster::agnes(biofam.OMsp, diss=T, method='gaverage')


hier.OMsp.W1 <- list()
hier.OMsp.A1 <- list()
hier.OMsp.S1 <- list()
hier.OMsp.C1 <- list()
hier.OMsp.WE1 <- list()
hier.OMsp.GA1 <- list()
for (k in 1:20) {
hier.OMsp.W1[[k]] <- cutree(hier.OMsp.W, k=k)
hier.OMsp.A1[[k]] <- cutree(hier.OMsp.A, k=k)
hier.OMsp.S1[[k]] <- cutree(hier.OMsp.S, k=k)
hier.OMsp.C1[[k]] <- cutree(hier.OMsp.C, k=k)
hier.OMsp.WE1[[k]] <- cutree(hier.OMsp.WE, k=k)
hier.OMsp.GA1[[k]] <- cutree(hier.OMsp.GA, k=k)
}

```

## K-medoids clustering

```{r}
kmedoids.OM <- list()
kmedoids.DHD <- list()
kmedoids.OMl <- list()
kmedoids.OMsl <- list()
kmedoids.OMsp <- list()
kmedoids.H <- list()
kmedoids.EU <-list()
kmedoids.CH <- list()
kmedoids.LS <- list()
kmedoids.LP <- list()
kmedoids.RL <- list()
kmedoids.N <- list()
kmedoids.NT <- list()
kmedoids.SV <- list()
for (k in 1:20) {
kmedoids.OM[[k]] <- cluster::pam(biofam.OM, diss=T, k=k)
kmedoids.DHD[[k]] <- cluster::pam(biofam.DHD, diss=T, k=k)
kmedoids.OMl[[k]] <- cluster::pam(biofam.OMl, diss=T, k=k)
kmedoids.OMsl[[k]] <- cluster::pam(biofam.OMsl, diss=T, k=k)
kmedoids.OMsp[[k]] <- cluster::pam(biofam.OMsp, diss=T, k=k)
kmedoids.H[[k]] <- cluster::pam(biofam.H, diss=T, k=k)
kmedoids.EU[[k]] <- cluster::pam(biofam.EU, diss=T, k=k)
kmedoids.CH[[k]] <- cluster::pam(biofam.CH, diss=T, k=k)
kmedoids.LS[[k]] <- cluster::pam(biofam.LS, diss=T, k=k)
kmedoids.LP[[k]] <- cluster::pam(biofam.LP, diss=T, k=k)
kmedoids.RL[[k]] <- cluster::pam(biofam.RL, diss=T, k=k)
kmedoids.N[[k]] <- cluster::pam(biofam.N, diss=T, k=k)
kmedoids.NT[[k]] <- cluster::pam(biofam.NT, diss=T, k=k)
kmedoids.SV[[k]] <- cluster::pam(biofam.SV, diss=T, k=k)
}





```



## State distribution plots
### Hierachical clusters
```{r}

seqplot(biofam.seq, type='d', group=hier.DHD.W1)

seqplot(biofam.seq, type='d', group=hier.DHD.A1)

seqplot(biofam.seq, type='d', group=hier.DHD.S1)

seqplot(biofam.seq, type='d', group=hier.DHD.C1)

seqplot(biofam.seq, type='d', group=hier.DHD.WE1)

seqplot(biofam.seq, type='d', group=hier.OM.GA1)

seqplot(biofam.seq, type='d', group=hier.OM.A1)

seqplot(biofam.seq, type='d', group=hier.OM.S1)

seqplot(biofam.seq, type='d', group=hier.OM.C1)

seqplot(biofam.seq, type='d', group=hier.OM.WE1)

seqplot(biofam.seq, type='d', group=hier.OM.GA1)



```

## K-medoid clusters
```{r}
seqplot(biofam.seq, type='d', group=kmedoids.DHD$clustering)
seqplot(biofam.seq, type='d', group=kmedoids.EU$clustering)
seqplot(biofam.seq, type='d', group=kmedoids.OMl$clustering)
seqplot(biofam.seq, type='d', group=kmedoids.OMsl$clustering)
seqplot(biofam.seq, type='d', group=kmedoids.OMsp$clustering)
seqplot(biofam.seq, type='d', group=kmedoids.H$clustering)
seqplot(biofam.seq, type='d', group=kmedoids.CH$clustering)
seqplot(biofam.seq, type='d', group=kmedoids.EU$clustering)
seqplot(biofam.seq, type='d', group=kmedoids.CH$clustering)
seqplot(biofam.seq, type='d', group=kmedoids.LS$clustering)
seqplot(biofam.seq, type='d', group=kmedoids.LP$clustering)
seqplot(biofam.seq, type='d', group=kmedoids.RL$clustering)
seqplot(biofam.seq, type='d', group=kmedoids.N$clustering)
seqplot(biofam.seq, type='d', group=kmedoids.NT$clustering)

```
## model with different clustering methods
```{r}
fit <- multinom(kmedoids.DHD$clustering ~ ., biofam.cov, trace=F)
coef(fit)


```



### Checking the quality of the clustering solution
#### kmedoids
##### Average Silhouette width - plots
```{r}
## OM
silhouette <- rep(NA, 20-1)
for(k in 1:19) {
  silhouette[k] <- wcClusterQuality(biofam.OM, kmedoids.OM[[k+1]]$clustering)$stats["ASW"]
}
plot(silhouette, type="l", xaxt="n")
axis(1, at=1:19, labels=2:20)


## DHD
silhouette <- rep(NA, 20-1)
for(k in 1:19) {
  silhouette[k] <- wcClusterQuality(biofam.DHD, kmedoids.DHD[[k+1]]$clustering)$stats["ASW"]
}
plot(silhouette, type="l", xaxt="n")
axis(1, at=1:19, labels=2:20)

## CH
silhouette <- rep(NA, 20-1)
for(k in 1:19) {
  silhouette[k] <- wcClusterQuality(biofam.CH, kmedoids.CH[[k+1]]$clustering)$stats["ASW"]
}
plot(silhouette, type="l", xaxt="n")
axis(1, at=1:19, labels=2:20)

## EU
silhouette <- rep(NA, 20-1)
for(k in 1:19) {
  silhouette[k] <- wcClusterQuality(biofam.EU, kmedoids.EU[[k+1]]$clustering)$stats["ASW"]
}
plot(silhouette, type="l", xaxt="n")
axis(1, at=1:19, labels=2:20)

## H
silhouette <- rep(NA, 20-1)
for(k in 1:19) {
  silhouette[k] <- wcClusterQuality(biofam.H, kmedoids.H[[k+1]]$clustering)$stats["ASW"]
}
plot(silhouette, type="l", xaxt="n")
axis(1, at=1:19, labels=2:20)

## LP
silhouette <- rep(NA, 20-1)
for(k in 1:19) {
  silhouette[k] <- wcClusterQuality(biofam.LP, kmedoids.LP[[k+1]]$clustering)$stats["ASW"]
}
plot(silhouette, type="l", xaxt="n")
axis(1, at=1:19, labels=2:20)

## LS
silhouette <- rep(NA, 20-1)
for(k in 1:19) {
  silhouette[k] <- wcClusterQuality(biofam.LS, kmedoids.LS[[k+1]]$clustering)$stats["ASW"]
}
plot(silhouette, type="l", xaxt="n")
axis(1, at=1:19, labels=2:20)

## N
silhouette <- rep(NA, 20-1)
for(k in 1:19) {
  silhouette[k] <- wcClusterQuality(biofam.N, kmedoids.N[[k+1]]$clustering)$stats["ASW"]
}
plot(silhouette, type="l", xaxt="n")
axis(1, at=1:19, labels=2:20)

## NT
silhouette <- rep(NA, 20-1)
for(k in 1:19) {
  silhouette[k] <- wcClusterQuality(biofam.NT, kmedoids.NT[[k+1]]$clustering)$stats["ASW"]
}
plot(silhouette, type="l", xaxt="n")
axis(1, at=1:19, labels=2:20)

## OMl
silhouette <- rep(NA, 20-1)
for(k in 1:19) {
  silhouette[k] <- wcClusterQuality(biofam.OMl, kmedoids.OMl[[k+1]]$clustering)$stats["ASW"]
}
plot(silhouette, type="l", xaxt="n")
axis(1, at=1:19, labels=2:20)

## OMsl
silhouette <- rep(NA, 20-1)
for(k in 1:19) {
  silhouette[k] <- wcClusterQuality(biofam.OMsl, kmedoids.OMsl[[k+1]]$clustering)$stats["ASW"]
}
plot(silhouette, type="l", xaxt="n")
axis(1, at=1:19, labels=2:20)

## OMsp
silhouette <- rep(NA, 20-1)
for(k in 1:19) {
  silhouette[k] <- wcClusterQuality(biofam.OMsp, kmedoids.OMsp[[k+1]]$clustering)$stats["ASW"]
}
plot(silhouette, type="l", xaxt="n")
axis(1, at=1:19, labels=2:20)

## RL
silhouette <- rep(NA, 20-1)
for(k in 1:19) {
  silhouette[k] <- wcClusterQuality(biofam.RL, kmedoids.RL[[k+1]]$clustering)$stats["ASW"]
}
plot(silhouette, type="l", xaxt="n")
axis(1, at=1:19, labels=2:20)

```


##### Quality Measures
```{r}
## CH
pamRange <- wcKMedRange(biofam.CH, kvals=2:20, weights=biofam$weights)
summary(pamRange, max.rank=2)

## DHD
pamRange1 <- wcKMedRange(biofam.DHD, kvals=2:20, weights=biofam$weights)
summary(pamRange1, max.rank=2)

## EU
pamRange2 <- wcKMedRange(biofam.EU, kvals=2:20, weights=biofam$weights)
summary(pamRange2, max.rank=2)

## H
pamRange3 <- wcKMedRange(biofam.H, kvals=2:20, weights=biofam$weights)
summary(pamRange3, max.rank=2)

## LP
pamRange4 <- wcKMedRange(biofam.LP, kvals=2:20, weights=biofam$weights)
summary(pamRange4, max.rank=2)

## LS
pamRange5 <- wcKMedRange(biofam.LS, kvals=2:20, weights=biofam$weights)
summary(pamRange5, max.rank=2)

## N
pamRange6 <- wcKMedRange(biofam.N, kvals=2:20, weights=biofam$weights)
summary(pamRange6, max.rank=2)

## NT
pamRange7 <- wcKMedRange(biofam.NT, kvals=2:20, weights=biofam$weights)
summary(pamRange7, max.rank=2)

## OM
pamRange8 <- wcKMedRange(biofam.OM, kvals=2:20, weights=biofam$weights)
summary(pamRange8, max.rank=2)

## OMl
pamRange9 <- wcKMedRange(biofam.OMl, kvals=2:20, weights=biofam$weights)
summary(pamRange9, max.rank=2)

## OMsl
pamRange10 <- wcKMedRange(biofam.OMsl, kvals=2:20, weights=biofam$weights)
summary(pamRange10, max.rank=2)

## OMsp
pamRange11 <- wcKMedRange(biofam.OMsp, kvals=2:20, weights=biofam$weights)
summary(pamRange11, max.rank=2)

## RL
pamRange12 <- wcKMedRange(biofam.RL, kvals=2:20, weights=biofam$weights)
summary(pamRange12, max.rank=2)



```

##### Finding the best clustering solution for each model
```{r}
## CH
which.max(pamRange$stats$ASW)
max(pamRange$stats$ASW)

## DHD
which.max(pamRange1$stats$ASW)
max(pamRange1$stats$ASW)

## EU
which.max(pamRange2$stats$ASW)
max(pamRange2$stats$ASW)

## H
which.max(pamRange3$stats$ASW)
max(pamRange3$stats$ASW)

## LP
which.max(pamRange4$stats$ASW)
max(pamRange4$stats$ASW)

## LS
which.max(pamRange5$stats$ASW)
max(pamRange5$stats$ASW)

## N
which.max(pamRange6$stats$ASW)
max(pamRange6$stats$ASW)

## NT
which.max(pamRange7$stats$ASW)
max(pamRange7$stats$ASW)

## OM
which.max(pamRange8$stats$ASW)
max(pamRange8$stats$ASW)

## OMl
which.max(pamRange9$stats$ASW)
max(pamRange9$stats$ASW)

## OMsl
which.max(pamRange10$stats$ASW)
max(pamRange10$stats$ASW)

## OMsp
which.max(pamRange11$stats$ASW)
max(pamRange11$stats$ASW)

## RL
which.max(pamRange12$stats$ASW)
max(pamRange12$stats$ASW)


```


#### hierarchical
##### OM 
```{r}
## average
clustQual <- as.clustrange(hier.OM.GA1, diss = biofam.OM, weights=biofam$weights, ncluster=20)


```




